---
title: "ETC3250/5250 - Project 2022"
author: Utkarsh Ahuja
output:
  html_document:
    df_print: paged
---

# **Introduction**

> In this report, we will be examining the data set *Celltype from gene motifs in mouse development*. This report comprises of tables, plots and explanations to grasp a better understanding of the provided data. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval = FALSE,
  message = FALSE,
  warning = FALSE)

library(readr)
library(dplyr)
library(kableExtra)
library(tidyverse)
library(tidymodels)
library(randomForest)
library(ggplot2)
library(naniar)
library(spinifex)
library(ggthemes)
library(ggridges)
```

```{r message=FALSE}
# Read data
tr <- read_csv("mouse_tr.csv") %>%
  mutate(celltype = factor(celltype))
ts <- read_csv("mouse_ts_mask.csv")
ts_sample<- read_csv(here::here("mouse_ts_samp.csv"))

```

# **Data pre-processing**


```{r}

# Count
tr$celltype %>% plyr::count()

# Plot
ggplot(tr) + 
  geom_bar(aes(y = celltype))+
  ggtitle ("Celltypes")+
  theme_solarized()


```

### Checking for duplicate/missing values

```{r}

pct_miss(tr)
n_miss(tr)

```

> After running the pct_miss() and n_miss() functions from the naniar package, we can conclude that there are no missing values in the given data set.

### Examining the data set and making necessary changes for modelling


```{r}
# merging data sets to add celltypes to the test set
ts<- merge(x=ts,y=ts_sample, by="location", all.x = TRUE)

```

> The first step in preparing our data for modelling was to combine the two data sets-ts and ts sample-into the ts data frame, where the celltype column was included along with the location and other fields.


### Conversion of labels in the data set

```{r}

# Conversion of the labels for training set

trainData_label <- as.integer(tr$celltype)-1

# setting null value to the column in tr

tr$celltype <- NULL

# Conversion of the labels for test set

testData_label <- as.integer(factor(ts$celltype))-1

# setting null value to the column in ts

ts$celltype <- NULL

```

> Because XGBoost only takes numeric vectors and classes in integer format, they will all begin with 0. As a result, the categorical class "celltype" will be transformed to integer format.


### Creation of training and testing data sets for the model

```{r}
# Creating the training and testing sets for the model

train_data <-  as.matrix(tr[,-1])

test_data <-  as.matrix(ts[,-1])

```

> The training data set is utilised to fit the model, whereas the testing data set is kept for validation purposes. This allows us to validate our model's performance (also known as "hold one out cross-validation").


### Creation of the xgb.DMatrix objects

```{r}
# Creating the xgb.DMatrix objects

dtrain <-  xgb.DMatrix(data=train_data, label= trainData_label)
dtest <-  xgb.DMatrix(data=test_data, label= testData_label)


```

> The training and testing data sets were then turned into xgb.DMatrix objects, which are utilised to fit the XGBoost model and predict future outcomes.


# **Summary of important variables**


```{r}

options(scipen = 999)

mouse_rf <- randomForest(celltype~.,data=tr[,-1], mtry=778, importance=TRUE,ntree=800)

mouse_rf$importance 

importance(mouse_rf, type = 2)

```

> RandomForest was utilised to analyse the data set and generate the best predictions before applying XGBoost for modelling. The mean decrease accuracy and mean decrease Gini score were used to determine the relevance of the variables for this class. I also examined the Variable Importance Plot to assess the significance of variables, but there was still no concrete evidence to determine which variables are essential and which are not. As a result, all variables were incorporated in the modelling. 


# **Model performance**

```{r eval=FALSE}

mouse_rf <- randomForest(celltype~.,data=tr[,-1], mtry=778, importance=TRUE,ntree=800)


```

> First, I used the randomForest classification algorithm with the best estimated mtry value of 778 and 800 trees. I was able to outperform the standard randomForest fit accuracy of 0.44/0.43. However, the XGBoost approach produced superior results.


```{r eval=FALSE}

# Fitting the xgboost model with cv = 10

mouse_xg <- train(celltype ~., data = tr[,-1],
                  method= "xgbTree",
                  trControl= trainControl("cv", number = 10))

```

```{r eval=FALSE}
# Determining the best tune for the model

mouse_xg$bestTune

```

```{r}
# Defining the main parameters for the classification

num_class = length(levels(celltype))
params = list(
  booster="gbtree",
  eta=0.3,
  max_depth=1,
  gamma=0,
  subsample=0.75,
  colsample_bytree=0.8,
  objective="multi:softprob",
  eval_metric="mlogloss",
  num_class = num_class
)

```

```{r}
# Training the model for classification

dfit <- xgb.train(
  params=params,
  data= dtrain,
  nrounds=150,
  nthreads=1,
  watchlist=list(train=dtrain, test=dtest),
  verbose=0
)

# Examine the model fit

dfit

 
```


> The best tune was established by fitting the XGBoost algorithm, which was afterwards utilised for a set of parameters that I discovered to be optimum using simple cross-validation. Later, we train the model and predict the likelihood of each observation in test data of being each celltype. 



# **References**

XGBoost Parameters — xgboost 1.6.1 documentation. (2022). Retrieved 27 May 2022, from https://xgboost.readthedocs.io/en/stable/parameter.html

Sarafian, R. (2022). Tree based models. Retrieved 27 May 2022, from
https://rstudio-pubs-static.s3.amazonaws.com/408996_debd530b72c844db9be5ef6ef9febc54.html

Kube, D. (2022). XGBoost Multinomial Classification Iris Example in R. Retrieved 27 May 2022, from https://rstudio-pubs-static.s3.amazonaws.com/456044_9c275b0718a64e6286751bb7c60ae42a.html

Malshe, A. (2022). 2.4 XGBoost | Data Analytics Applications. Retrieved 27 May 2022, from https://ashgreat.github.io/analyticsAppBook/xgboost



